\chapter{Conclusions and Future Work}

\epigraph{I'm completely cricketed out. If I never have to write another word about cricket again, I'll be a happy man.}{Joseph O'Neill}

\section{Conclusions}
We presented a Neural Network model for predicting cricket scores based on patterns in the runrate of a game. Neural Networks have been used in cricket for things such as 
generating commentary \cite{kumar2} or classifting shots based on footage \cite{foysal}, but we couldn't find a paper using them in this way before. 
The Neural Network model we built for predicting cricket scores did fairly well when given a full dataset, a $94\%$ corrolation with actual values is certainly not 
something to be glossed over. However, the aim of this project was for predicting results when the full data is not available due to an innings being cut short. 
In this domain, the network did considerably worse- correlating poorly with test data when monte-carlo methods were used to fill in the missing games, and only achieving only $50\%$ 
corrolation when the gaps were left as 0s. Despite this, applying the network to an actual tournament gave similar results to the standard method used for achieving the same task, and had 
marginal affect on the tournament's outcome. 


\section{Future Work}
Rather than decrasing/increasing scores via a proportion, it would be benificial to incorperate the fall of wicket densities looked at earlier in the paper in some way. This is one 
area our model doesn't incorperate at all, let alone well. Yet, it is a key aspect of cricket, and so in the future it could be highly effective if taken into account. \\

Using team specific data rather than an average of all games could be interesting thing to look into, but the computational cost of this would be high, and the accessability of the 
method would be poor as a result, defeating the point of making this a system easily employable across cricket. \\

As mentioned before, the lack of data available is a shame, but this is something that can't be helped due to the nature of the cricketing calender. It would be interesting to revisit this topic in a couple of years when more games have been played to see how results differ. One other avenue would be to look at combining the international games that have been included with domestic 50 over matches from around the world. At present, this would have increased the number of games by 436. \\ 

In appendix C, we discuss how the code used for this project is turned into a publicly available R package. In future, this will allow for easily experimenting with different neural 
network parameters such as the activation function and hidden layer arrangements. It goes without saying that this lends itself well to future work at building a much better neural network for predicting cricket scores.
Naturally, a lot of this is trial and error and so it would require a considerable amount of computing time to find an acceptable one. \\

In Section~\ref{lassoSec}, we saw how the order of ``most important'' variables was different, and so we could use the package to change the formula used by the network to take advantage of the more important overs only. Again this raises several more variables, since we would need to experiment with different numbers of important overs and such.
