\documentclass[11pt]{report}
\usepackage[utf8]{inputenc}
\usepackage[a4paper, total={6in,8in}, portrait, margin=1in]{geometry}
\usepackage{amssymb}
\usepackage{array}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{float}
\usepackage{mathtools}
\usepackage{listings}
\usepackage{xcolor}
\usepackage[misc]{ifsym}
\usepackage{indentfirst} 
\usepackage{amsthm}
\usepackage{appendix}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{fancyhdr}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{tikz}
\usepackage{dirtree}
\usepackage{epigraph}
\usetikzlibrary{positioning}
\pagestyle{fancy}


%Headers and Footers, thanks to Overleaf
%https://www.overleaf.com/learn/latex/How_to_Write_a_Thesis_in_LaTeX_(Part_2):_Page_Layout
\fancyhead{}
\fancyhead[RO,LE]{\small Improving Duckworth-Lewis: Statistical Methods for Revising Score Targets in Limited-Overs Cricket}
\fancyfoot{}
\fancyfoot[LE,RO]{\thepage}
\fancyfoot[LO,CE]{Chapter \thechapter}
\fancyfoot[CO,RE]{Matthew Knowles}

\DeclarePairedDelimiter{\ceil}{\lceil}{\rceil}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{remark}[theorem]{Remark}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{lemma}[theorem]{Lemma}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}

\lstset{style=mystyle}

\begin{document}

\input{titlepage}

\section*{Abstract}
A Neural Network model for predicting scores in the sport of Cricket is presented. The aim is to use our model to improve upon the results of the method currently employed by 
the International Cricket Council (ICC). We explore this current method, known as the DLS method \cite{stern} in 
some detail before going on to develop our own model. This is followed by some exploratory data analysis, which is imperative in making certain assumptions that aid 
in the building of the Neural Network. For the Neural Network, we take an in-depth look at the mathematical underpinning of building and training such a model. This 
is accompanied by code snippets demonstrating how these methods are implemented in a high-level programming language. We then perform an analysis of the results 
of our model before applying it to the data from the ICC 2019 Cricket World Cup, in order to see how it faired against the DLS. 
We find that our method does not predict scores perfectly, but does produce slightly more attainable score targets than DLS- as illustrated by the World Cup simulation 
study. 

\section*{Acknowledgements}
The author would like to thank Dr. Jessica Hargreaves for their invaluable supervision, encouragement and advice throughout this project. In addition,
many thanks to Michael Najdan at Kent County Cricket Club for their insight into how data is used in the professional game of cricket, and to Harrison Allen at Yorkshire County Cricket Club for demonstrating how limited-overs data was put into practice in preperation for games. 
I couldn't write a dissertation about cricket without mentioning the University of York Men's Cricket Club- it has been wonderful to share the playing field with you all over the past 4 years.
Finally, I would like to dedicate this dissertation to my friends from Lindley G; thank you for being the most wonderful and supportive people I've ever met. You'll never know how much you mean to me.

\section*{Statment of Originallity}
The enclosed content is my own work. Any work taken from another source has been appropriately acknowledged and referenced. 

\section*{Notes}
All code written in support of this project can be found on GitHub at: \\
\begin{itemize}
    \item \textit{https://github.com/mattknowles314/mmathThesis}. 
    \item \textit{https://github.com/mattknowles314/CricNet}.
\end{itemize}


\setcounter{tocdepth}{3}
\tableofcontents
\setcounter{tocdepth}{1}
\listoffigures


\input{Chapters/Intro}

\input{Chapters/dataChap}

\input{Chapters/DLSChap}

\input{Chapters/expData.tex}

\input{Chapters/SPRChap}

\input{Chapters/analysis}

\input{Chapters/wc2019sim}

\input{Chapters/conclusions.tex}

\appendix

\input{Chapters/qq.tex}

\input{Chapters/net.tex}

\input{Chapters/package.tex}

\bibliographystyle{apalike}
\bibliography{refs}{}


\end{document}
